---
title: "Case Study 2:  DDSAnalytics"
author: "Kristi Herman"
date: "03/11/2020"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Youtube Link
*https://youtu.be/B3NS19GLUCk 

## Introduction

* The purpose of this analysis is to determine that factors that lead to employee attrition, predict employee turnover, and predict monthly income.

```{r}
# Import dependencies
library(tidyverse)
library(dplyr)
library(ggplot2)
library(naniar)
library(class)
library(caret)
library(corrplot)
library(ggthemes)
library(ggpubr)
library(gplots)
library(arsenal)
library(mltools)
library(GGally)
library(e1071)

# Import employee data and prediction data
empl_df <-  read.csv("CaseStudy2-data.csv")

```

* Basic Data Exploration

```{r}

# Check to make sure data imported
dim(empl_df)
View(empl_df)

# Check to see if there's any missing data
gg_miss_var(empl_df)

# Get summary statistics
summary(empl_df)
str(empl_df)

# Confirm that these fields only have one value unique values, and min/max values are
unique(empl_df$Over18)
unique(empl_df$StandardHours)
unique(empl_df$EmployeeCount)

# Look at values in more fields
unique(empl_df$JobInvolvement)
str(unique(empl_df$EmployeeNumber))


```

* Data cleaning

```{r}

# Reorder columns
empl_data <- empl_df[c(1,11,3,2,4:10,12:36)]
head(empl_data)
str(empl_data)
dim(empl_data)

# Max and mins
min(empl_data[,36])
max(empl_data[,36])

# Drop Over18, StandardHours, EmployeeCount, ID, MonthlyRate
empl_data <- empl_data[ , -which(names(empl_data) %in% c("EmployeeNumber", "Over18","StandardHours", "EmployeeCount", "MonthlyRate", "HourlyRate", "DailyRate"))]

head(empl_data)

```

* Overall Attrition for the Dataset is 16%

```{r}

# Determine and Plot Attrition for the Dataset
attr <- empl_data %>% group_by(Attrition) %>% summarize(count=n())
attr_pct <- attr$count[2]/dim(empl_data)[1]*100

p <- attr %>% ggplot(aes(x = Attrition, y = count)) +
  geom_bar(stat = "identity", fill = "blue", alpha = .6) +
  ggtitle("Attrition Overview") + xlab("Attrition") + ylab("Number of Employees") + 
  geom_text(aes(y = count, label = count), fontface = "bold", vjust = 1.2, color = "white", size = 3)  

p

```

* Average Monthly Salary By Job Role - Managers & Research Directors have the highest average monthly salaries

```{r}

#  Get average monthly salary for each job satisfaction categories
income <- empl_data %>% group_by(JobRole) %>% summarize(mean_size = round(mean(MonthlyIncome)), digits = 2)

# Plot data
p <- income %>%
  ggplot(aes(x = reorder(JobRole, -mean_size), y = mean_size)) +
  geom_bar(stat = "identity", fill = "blue", alpha = .6) +
  ggtitle("Job Role & Average Monthly Salary") + xlab("Job Role") + ylab("Average Monthly Salary") + 
  geom_text(aes(y = mean_size, label = mean_size), fontface = "bold", vjust = 1.2, color = "white", size = 3)  

p + rotate_x_text(45)

```

* Job Satisfaction By Job Role

```{r}

#  Get average monthly salary for each job satisfaction categories
satis <- empl_data %>% group_by(JobRole) 

# Distribution of job satisfaction by job role (Scatter)
p <- ggplot(data = empl_data) +
  geom_point(mapping = aes(x = JobRole, y = JobSatisfaction, color = JobSatisfaction), position ="jitter") +
  ggtitle("Job Satisfaction By Job Role") + xlab("Job Role") + ylab("Job Satisfaction")

p + rotate_x_text(45)


```

* Average Job Satisfaction By Job Role - Managers & Research Directors have the lowest average job satisfaction

```{r}

#  Get average monthly salary for each job satisfaction categories
satis <- empl_data %>% group_by(JobRole) %>% summarize(mean_size = round(mean(JobSatisfaction), digits = 2))

# Plot data
p <- satis %>%
  ggplot(aes(x = reorder(JobRole, mean_size), y = mean_size)) +
  geom_bar(stat = "identity", fill = "blue", alpha = .6) +
  ggtitle("Job Role & Average Job Satisfaction") + xlab("Job Role") + ylab("Average Job Satisfaction")  +
  geom_text(aes(y = mean_size, label = mean_size), fontface = "bold", vjust = 1.2, color = "white", size = 3) 

p + rotate_x_text(45)



```


* Average Job Involvement By Job Role - Sales Reps have the Lowest Average Job Involvement

```{r}

#  Get average monthly salary for each job satisfaction categories
involve <- empl_data %>% group_by(JobRole) %>% summarize(mean_size = round(mean(JobInvolvement), digits = 2))

# Plot data
p <- involve %>%
  ggplot(aes(x = reorder(JobRole, mean_size), y = mean_size)) +
  geom_bar(stat = "identity", fill = "blue", alpha = .6) +
  ggtitle("Job Role & Average Job Involvement") + xlab("Job Role") + ylab("Average Job Involvement")  +
  geom_text(aes(y = mean_size, label = mean_size), fontface = "bold", vjust = 1.2, color = "white", size = 3) 

p + rotate_x_text(45)



```

* Attrition By Overtime - Non-Salaried or Overtime Employees have Higher Attrition

```{r}

#  Group jobs by job role and calculate percentage of attrition
overtime <- empl_data %>% group_by(OverTime, Attrition) %>% summarize(count=n()) %>% mutate(total_by_overtime = sum(count)) 
overtime$pct <- round(overtime$count/overtime$total_by_overtime, digits = 2)*100

# Plot attrition by job role

p <- ggplot(data=overtime, aes(x=OverTime, y=pct, fill=Attrition)) + geom_bar(stat="identity", alpha = .6) +
  ggtitle("Attrition by Overtime") + xlab("Overtime") + ylab("Attrition (%)") + scale_fill_manual(values = c("blue","orange")) +
  geom_text(aes(y = pct, label = pct), fontface = "bold", nudge_y = -1.5, color = "white", size = 3)

p

```

* Attrition By Marital Status - Singles have the Highest Attrition While Divorced Employees have the Lowest Attrition

```{r}

#  Group jobs by job role and calculate percentage of attrition
married <- empl_data %>% group_by(MaritalStatus, Attrition) %>% summarize(count=n()) %>% mutate(total_by_marriage = sum(count)) 
married$pct <- round(married$count/married$total_by_marriage, digits = 2)*100

# Plot attrition by job role

p <- ggplot(data=married, aes(x=MaritalStatus, y=pct, fill=Attrition)) + geom_bar(stat="identity", alpha = .6) +
  ggtitle("Attrition by Marital Status") + xlab("Marital Status") + ylab("Attrition (%)") + scale_fill_manual(values = c("blue","orange"))  +
  geom_text(aes(y = pct, label = pct), fontface = "bold", vjust = 1, color = "white", size = 3)

p + rotate_x_text(45)

```

* Attrition By Job Involvement - The Higher the Job Involvement, the Lower the Attrition is

```{r}

#  Group jobs by job role and calculate percentage of attrition
job_involve <- empl_data %>% group_by(JobInvolvement, Attrition) %>% summarize(count=n()) %>% mutate(total_by_involvement = sum(count)) 
job_involve$pct <- round(job_involve$count/job_involve$total_by_involvement, digits = 2)*100

# Plot attrition by job role

p <- ggplot(data=job_involve, aes(x=JobInvolvement, y=pct, fill=Attrition)) + geom_bar(stat="identity", alpha = .6) +
  ggtitle("Attrition by Job Involvement") + xlab("Job Involvement") + ylab("Attrition (%)")  + scale_fill_manual(values = c("blue","orange")) + 
  geom_text(aes(y = pct, label = pct), fontface = "bold", vjust = 1, color = "white", size = 3) 

p

```

* Attrition By Role - Sales Reps have the Highest Attrition, While Research Directors, Manufacturing Directors, and Managers have the Lowest Attrition

```{r}

#  Group jobs by job role and calculate percentage of attrition
job_role <- empl_data %>% group_by(JobRole, Attrition) %>% summarize(count=n()) %>% mutate(total_by_role = sum(count)) 
job_role$pct <- round(job_role$count/job_role$total_by_role, digits = 2)*100

# Plot attrition by job role

p <- ggplot(data=job_role, aes(x=JobRole, y=pct, fill=Attrition)) + 
  geom_bar(stat="identity", alpha = .6) +
  ggtitle("Attrition by Role") + xlab("Job Role") + ylab("Attrition (%)") + scale_fill_manual(values = c("blue","orange")) + 
geom_text(aes(y = pct, label = pct), fontface = "bold", vjust = 1, color = "white", size = 3)

p + rotate_x_text(45)

```

* Attrition By Salary - Higher Paying Jobs Have Lower Attrition

```{r}
income_attr <- empl_data 
income_attr$bin <- bin_data(income_attr$MonthlyIncome, bins=c(0, 4000, 8000, 12000, 16000, 20000), boundaryType = "[lorc")

income_attr <- income_attr %>% group_by(bin, Attrition) %>% summarize(count=n()) %>% mutate(total_by_bin = sum(count)) 
income_attr$pct <- round(income_attr$count/income_attr$total_by_bin, digits = 2)*100

p <- ggplot(data=income_attr, aes(x=bin, y=pct, fill=Attrition)) + geom_bar(stat="identity", alpha = .6) +
  ggtitle("Attrition by Salary") + xlab("Salary") + ylab("Attrition (%)")  + scale_fill_manual(values = c("blue","orange")) + 
  geom_text(aes(y = pct, label = pct), fontface = "bold", vjust = 1, color = "white", size = 3) 

p


```


* Attrition By Job Satisfaction - Higher Job Satisfaction Does Lead to Lower Attrition Overall

```{r}

#  Group jobs by job role and calculate percentage of attrition
job_satis <- empl_data %>% group_by(JobSatisfaction, Attrition) %>% summarize(count=n()) %>% mutate(total_by_satis = sum(count)) 
job_satis$pct <- round(job_satis$count/job_satis$total_by_satis, digits = 2)*100

# Plot attrition by job role

p <- ggplot(data=job_satis, aes(x=JobSatisfaction, y=pct, fill=Attrition)) + geom_bar(stat="identity", alpha = .6) +
  ggtitle("Attrition by Job Satisfaction") + xlab("Job Satisfaction") + ylab("Attrition (%)")  + scale_fill_manual(values = c("blue","orange")) + 
  geom_text(aes(y = pct, label = pct), fontface = "bold", vjust = 1, color = "white", size = 3) 

p

```

* Overtime and Job Involvement - Non-Salaried, Overtime Eligible Employees have Lower Job Involvement

```{r}

#  Group jobs by job role and calculate percentage of attrition
job_over <- empl_data %>% group_by(JobInvolvement, OverTime) %>% summarize(count=n()) %>% mutate(total_by_involvement = sum(count)) 
job_over$pct <- round(job_over$count/job_over$total_by_involvement, digits = 2)*100

# Plot attrition by job role

p <- ggplot(data=job_over, aes(x=JobInvolvement, y=pct, fill=OverTime)) + geom_bar(stat="identity", alpha = .6) +
  ggtitle("Overtime by Job Involvement") + xlab("Job Involvement") + ylab("Overtime(%)")  + scale_fill_manual(values = c("blue","orange")) + 
  geom_text(aes(y = pct, label = pct), fontface = "bold", vjust = 1, color = "white", size = 3) 

p

```

* Marital Status and Job Involvement

```{r}

#  Group jobs by job role and calculate percentage of attrition
marital_status <- empl_data %>% group_by(JobInvolvement, MaritalStatus) %>% summarize(count=n()) %>% mutate(total_by_involvement = sum(count)) 
marital_status$pct <- round(marital_status$count/marital_status$total_by_involvement, digits = 2)*100

# Plot attrition by job role

p <- ggplot(data=marital_status, aes(x=JobInvolvement, y=pct, fill=MaritalStatus)) + geom_bar(stat="identity", alpha = .6) +
  ggtitle("Marital Status by Job Involvement") + xlab("Job Involvement") + ylab("Marital Status(%)")  + scale_fill_manual(values = c("blue","orange", "red")) 
p

```
* Exploration - facet_wrap on all columns

```{r}

p <- ggplot() + 
  geom_histogram(aes(x=JobInvolvement, fill = "r", color="r"), alpha=.7, data=empl_data, stat = "bin") +
  ggtitle("Job Involvement") + xlab("Job Involvement") + ylab("Number of Employees") +
  scale_color_manual(name="group", values=c("r" = "red"), labels=c("r"="Employees")) 

p

p <- ggplot() + 
  geom_histogram(aes(x=JobInvolvement), alpha=.7, data=empl_data, stat = "bin") +
  ggtitle("Job Involvement") + xlab("Job Involvement") + ylab("Number of Employees") +
  scale_color_manual(name="Employees", values=c("r" = "red")) 

p

```


* Exploration - facet_wrap on all columns

```{r, fig.width=15, fig.height=15}

# Get histograms of the data
empl_data %>% keep(is.numeric) %>% gather %>% ggplot(aes(x = value)) + facet_wrap(~key, scales = "free")+ geom_histogram() 

```

* Correlation plot

```{r, fig.width=15, fig.height=15}

empl_data_corr <- empl_data

# Make factors into integers
empl_data_corr[c("JobRole","BusinessTravel","Department","EducationField","Gender", "JobSatisfaction", "MaritalStatus", "OverTime")] <-  lapply(empl_data[c("JobRole","BusinessTravel","Department","EducationField","Gender", "JobSatisfaction", "MaritalStatus", "OverTime")], as.integer)

# Change Attrition to an integer
empl_data_corr$Attrition <- as.numeric(empl_data_corr$Attrition)

empl_data_corr %>%
  keep(is.numeric) %>%
  cor %>% 
  corrplot("upper", diag = FALSE, order = "FPC", tl.pos = "td", tl.cex = 1, method = "color", tl.col = "black", addCoef.col = "grey", number.digits = 2, title="Correlation of Numerical Attributes", mar=c(0,0,1,0))

# From this graph, not many of the attributes seem to be highly correlated with Attrition.  But these are the highest two attributes:  JobInvolvement, OverTime, and Marital Status.  

```


* GGpairs with the some of the features

```{r, fig.width=10, fig.height=5}
empl_data %>%
select(JobInvolvement, TotalWorkingYears, Attrition) %>%
ggpairs(mapping = aes(color = Attrition)) 
```


* KNN for each feature:  Conclusion - The specifity is not high enough with these variables.  There is not much variation when K is 3, 5, 30, or 100.

```{r}

# New dataset
empl_knn <- empl_data

# Make factors into integers
empl_knn[c("JobRole","BusinessTravel","Department","EducationField","Gender", "JobSatisfaction", "MaritalStatus", "OverTime")] <- lapply(empl_knn[c("JobRole","BusinessTravel","Department","EducationField","Gender", "JobSatisfaction", "MaritalStatus", "OverTime")], as.integer)

# Split training data into train and test data

set.seed(sample(1:100, 1))
split_perc = .75
train_idx = sample(1:dim(empl_knn)[1],round(split_perc * dim(empl_knn)[1])) 
train_idx
train = empl_knn[train_idx,]
test = empl_knn[-train_idx,]

# Run the classification model on all the features
classif_knn <-  knn(train[, 3:29],test[, 3:29],train$Attrition, prob = TRUE, k = 5)
results <- confusionMatrix(table(classif_knn,test$Attrition))
results


# Create new dataframes for accuracy results
acc_df_knn <- data.frame(feature_names=character(0),acc=numeric(0), sens=numeric(0), spec=numeric(0))

# New test/train data
set.seed(sample(1:100, 1))
split_perc = .75
train_idx = sample(1:dim(empl_knn)[1],round(split_perc * dim(empl_knn)[1])) 
train_idx
train_knn= empl_knn[train_idx,]
test_knn = empl_knn[-train_idx,]

# Run KNN on all of the features
for (i in 3:dim(empl_knn)[2]){ 
  if (i > 2) {
    classif_knn = knn(data.frame(train_knn[, i]),data.frame(test_knn[, i]),train_knn$Attrition, prob = TRUE, k = 100)
    results <- confusionMatrix(table(classif_knn,test_knn$Attrition))
    temp3 <- data.frame(feature_names=colnames(empl_data[i]), acc=results[[3]][1], sens = results$byClass[1], spec = results$byClass[2])
    acc_df_knn <- rbind(acc_df_knn, temp3)
  }
}

unique(acc_df_knn$acc)
unique(acc_df_knn$sens)
unique(acc_df_knn$spec)


```

* KNN Cross Validation (CV) for each feature

```{r}

# Create new dataframes for accuracy results
acc_df_knn_cv <- data.frame(feature_names=character(0),acc=numeric(0), sens=numeric(0), spec=numeric(0))

# Run KNN CV on all of the features
for (i in 3:dim(empl_knn)[2]){ 
  if (i > 2) {
    results_cv = confusionMatrix(table(knn.cv(empl_knn[,i], empl_knn$Attrition, k = 5), empl_knn$Attrition))
    temp4 <- data.frame(feature_names=colnames(empl_knn[i]), acc=results_cv[[3]][1], sens = results_cv$byClass[1], spec = results_cv$byClass[2])
    acc_df_knn_cv <- rbind(acc_df_knn_cv, temp4)
  }
}

unique(acc_df_knn_cv$acc)
unique(acc_df_knn_cv$sens)
unique(acc_df_knn_cv$spec)


# In this one TotalWorkingYears had the highest accuracy, but the specificity is still low


```

* Naive Bayes - Train Model

```{r}

# Create a new df
empl_data2 <- empl_data[c(1:29)]

# Split training data into train and test data
set.seed(sample(1:100, 1))
split_perc = .75
train_idx = sample(1:dim(empl_data2)[1],round(split_perc * dim(empl_data2)[1])) 
train_idx
train = empl_data2[train_idx,]
test = empl_data2[-train_idx,]

# Create new dataframes for accuracy results
acc_df_nb <- data.frame(feature_names=character(0),acc=numeric(0), sens=numeric(0), spec=numeric(0))

model = naiveBayes(train[,3:29], train$Attrition, laplace = 1)
results_nb = confusionMatrix(table(predict(model,test[,3:29]),test$Attrition))
results_nb

```

* Naive Bayes - Predict Attrition on new data

```{r}

empl_attr_prdt <-  read.csv("CaseStudy2CompSet_No_Attrition.csv")

# Add column for Attrition
empl_attr_prdt$Attrition <- NA

# Reorder columns
empl_attr_prdt <- empl_attr_prdt[c(1,36,2:35)]

# Drop Over18, StandardHours, EmployeeCount, ID, MonthlyRate
empl_attr_prdt <- empl_attr_prdt[ , -which(names(empl_attr_prdt) %in% c("EmployeeNumber", "Over18","StandardHours", "EmployeeCount", "MonthlyRate", "HourlyRate", "DailyRate"))]

# Compare datasets
summary(comparedf(empl_data2, empl_attr_prdt))

# Add prediction to MonthlyIncome column
empl_attr_prdt$Attrition <- model %>% predict(empl_attr_prdt)

# Export all data
write.csv(empl_attr_prdt, "Case2PredictionsHermanAttrition-Full.csv", row.names=FALSE)

# Export ID and Attrition
export_attr <- empl_attr_prdt[c(1,2)]
write.csv(export_attr, "Case2PredictionsHermanAttrition.csv", row.names=FALSE)

```

* Run NB on each feature individually

```{r}

# Run NB on all of the features
for (i in 3:dim(empl_data)[2]){ 
  if (i > 2) {
    model = naiveBayes(train[,i], train$Attrition, laplace = 1)
    results_nb = confusionMatrix(table(predict(model,test[,i]),test$Attrition))
    temp5 <- data.frame(acc = c(results_nb$overall[1]), sens = c(results_nb$byClass[1]), spec = c(results_nb$byClass[2]))
    acc_df_nb <-  rbind(acc_df_nb, temp5)
  }
}

str(test)

acc_df_nb
max(acc_df_nb$acc)
min(acc_df_nb$sens)
max(acc_df_nb$spec)


# Sort data
sorted_results_nb <- acc_df_nb[order(-acc_df_nb$acc),][1:20,]
sorted_results_nb

# All values are the same with this model with factors and integers
head(sorted_results_nb)

```

* Linear Regression - Train Model

```{r}

# Dataset to train on 
empl_data

# Create a training df with only numeric variables
train_df <- empl_data %>% keep(is.numeric)

# Split the data
set.seed(123)
tr_samples <- train_df$MonthlyIncome %>% createDataPartition(p = .95, list = FALSE)
tr  <- train_df[tr_samples, ]
td <- train_df[-tr_samples, ]

# Build the model, use monthly income 
model <- lm(MonthlyIncome ~., data = tr)

# Summarize the model
summary(model)

# Make predictions on the test dataset
preds <- model %>% predict(td)

# Evaluate performance of the model with RMSE
RMSE(preds, td$MonthlyIncome)

```

* Linear Regression - Predict Salary on new data
```{r}

# Import csv file without salary data
empl_salary_prdt <-  read.csv("CaseStudy2CompSet_No_Salary.csv")

# Change Attrition to an integer
empl_salary_prdt$Attrition <- as.numeric(empl_salary_prdt$Attrition)

# Drop Over18, StandardHours, EmployeeCount, ID, MonthlyRate
empl_salary_prdt <- empl_salary_prdt[ , -which(names(empl_salary_prdt) %in% c("EmployeeNumber", "Over18","StandardHours", "EmployeeCount", "MonthlyRate", "HourlyRate", "DailyRate"))]

# Fix ID name
colnames(empl_salary_prdt)[1] <- "ID"

# Add a blank column for monthly income
income_pred <- empl_salary_prdt %>% keep(is.numeric)
income_pred$MonthlyIncome <- NA 

# Add prediction to MonthlyIncome column
income_pred$MonthlyIncome <- model %>% predict(income_pred)
grep("MonthlyIncome", colnames(income_pred))

View(income_pred)

# Export full dataset
write.csv(income_pred, "Case2PredictionsHermanSalary-Full.csv", row.names=FALSE)

# Export ID and MonthlyIncome
export_df <- income_pred[c(1,22)]
write.csv(export_df, "Case2PredictionsHermanSalary.csv", row.names=FALSE)

```



